Figure Extra

### 2023-12-14 Figures/Figures_nonTCGA.Rmd
add Vanderbilt + try fix bug parse
given the complex naming of this cohort ground truth I have some
problems when cleaning the correlation table.

I try and fix this by changing processing of ground truth and add some
comments for ease of ussage

    results <- tibble(Samples = deconv_data[[key]])

@@

    for (method in unique(names_table[,1,drop=T])){

@@

          for (celltype in names(signatures_dictionary[[signature]])){
            if (length(signatures_dictionary[[signature]][[celltype]]) > 0){
              name_of_summary <- paste(to_match, celltype, sep = "_")
              columns_to_summarize <- paste(to_match, signatures_dictionary[[signature]][[celltype]], sep = ".")

@@

            deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% rowSums(.) -> results[name_of_summary]
        }
      }
    }
    }
    results %>% dplyr::select(-Samples) %>% dplyr::select(which(!colSums(., na.rm=TRUE) %in% 0)) %>% bind_cols(Samples = results$Samples,.)
}
################################################################################

@@

```{r}
benchmark_corrplots <- function(deconv_data, true_estimates, dict, key, title){
  # build the data
  true_subset_ <- true_estimates %>%  na.exclude() %>%
    arrange(factor(get(key), levels = get(key)))

@@

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = dict, key = key)
  identical(true_subset[[key]], deconv_subset_summary[[key]]) %>% stopifnot()

  to_corr <- inner_join(true_subset, deconv_subset_summary) %>% dplyr::select(-any_of(key))
  ct_corrplots <- list()
  store_corrs <- tibble(measure1 = character(),
                          measure2 = character(),
                          r = numeric(),
                          n = numeric(),
                          p = numeric(),
                          sig_p = logical(),
                          p_if_sig = numeric(),
                          r_if_sig = numeric())

    for (ct in unique(dplyr::select(true_estimates, -any_of(key)) %>% names())){

@@

      deconv_ct_names <- to_corr %>%
        # segregate by ct
        dplyr::select(matches(paste0(ct, "$"))) %>%
        formatted_cors(cor.stat = "pearson") %>%
        # reduce to corrs just to ground truth
        dplyr::filter(measure1 == all_of(ct), measure2 != all_of(ct)) %>%
        # rename ground truth as cancer type
        dplyr::mutate(measure2 = str_remove_all(measure2,  paste(substring_remove, collapse = "|")))
      #add to the rest of cancer type correlations
      store_corrs <- add_row(store_corrs , deconv_ct_names)
    }

@@

>>>>

  # Create a tibble to store equivalences
    results <- tibble(Samples = deconv_data[[key]])

@@

    # Iterate throughout methods and their signatures, saving with standard notation
    for (method in unique(names_table[,1,drop=T])){

@@

          for (celltype in names(signatures_dictionary[[signature]])){
            if (length(signatures_dictionary[[signature]][[celltype]]) > 0){
              name_of_summary <- paste(to_match, celltype, sep = "_") # New name
              columns_to_summarize <- paste(to_match, signatures_dictionary[[signature]][[celltype]], sep = ".") # Columns to sum of the deconv to have the new name

@@

            deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% rowSums(.) -> results[name_of_summary]
        }
      }
    }
    }
    # remove empty signatures
    results %>% dplyr::select(-Samples) %>% dplyr::select(which(!colSums(., na.rm=TRUE) %in% 0)) %>% bind_cols(Samples = results$Samples,.)
}
################################################################################

@@

```{r}
benchmark_corrplots <- function(deconv_data, true_estimates, dict, key, title){
  # build the data to parse all the estimations to ground truth
  true_subset_ <- true_estimates %>%  na.exclude() %>%
    arrange(factor(get(key), levels = get(key))) %>%
    dplyr:::rename_with(~str_replace(.,"_", ".")) %>%
    dplyr:::rename_with(~str_replace(.," ", "."))

@@

  # summarize those predictions constituting each ground truth celltype and standardize naming
  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = dict, key = key)
  identical(true_subset[[key]], deconv_subset_summary[[key]]) %>% stopifnot()

  # Correlate predictions to ground truth
  to_corr <- inner_join(true_subset, deconv_subset_summary) %>% dplyr::select(-any_of(key))
  ct_corrplots <- list()
  store_corrs <- tibble(measure1 = character(),
                          measure2 = character(),
                          r = numeric(),
                          n = numeric(),
                          p = numeric(),
                          sig_p = logical(),
                          p_if_sig = numeric(),
                          r_if_sig = numeric())

    for (ct in unique(dplyr::select(true_subset, -any_of(key)) %>% names())){

@@

      deconv_ct_names <- to_corr %>%
        # segregate by ct
        dplyr::select(matches(paste0(ct, "$"))) %>%
        formatted_cors(cor.stat = "pearson") %>%
        # reduce to just to ground truth on one side, deconv on the other
        dplyr::filter(measure1 == all_of(ct), measure2 != all_of(ct)) %>%
        # rename ground truth as cancer type
        dplyr::mutate(measure2 = str_remove_all(measure2,  paste(substring_remove, collapse = "|")))
      #add to the rest of cancer type correlations
      store_corrs <- add_row(store_corrs , deconv_ct_names)
    }

@@

# Vanderbilt
## Data loading
```{r}
deconv_Vanderbilt <- read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Data/deconvolution_for_gemdecan/Vanderbilt_08122023/all_deconvolutions_TPM_Counts_Van.txt") %>%
  dplyr::rename_with(~str_replace(.,"__", "_CBSX_")) %>%
  dplyr::rename_with(~str_replace_all(.,"\\.", "_")) %>%
  dplyr::rename_with(~str_replace_all(.,"__", "_"))#! fix naming of deconv results

truth_Vanderbilt <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Data/deconvolution_for_gemdecan/Vanderbilt_08122023/Cell_proportions.csv") %>%
  rename(Samples = "...1")

```{r}
# Lists of celltypes of each signature
## signature dictionary construction (can use . as regexp!)
## uncoment print of functions to check that you detect what you want.
## non matching signatures/methods (all 0) are removed.
dict_Vanderbilt <- list()
dict_Vanderbilt[["BPRNACan"]] <- list(`B cells` = c("B"),
                                      CD4 = c("CD4"),
                                      CD8 = c("CD8"),
                                      `Endothelial cells` = c(),
                                      Cancer = c("cancer"),
                                      Fibroblasts = c(),
                                      `M1 Macrophages` = c("M1"),
                                      `M2 macrophages` = c("M2"),
                                      `Mast cells` = c(),
                                      `Mural cells` = c(),
                                      NK = c("NK"),
                                      `Plasma B cells` = c(),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c(),
                                      `T reg. cells` = c()) # Missing Monocytes, M0 and Neutrophils

dict_Vanderbilt[["BPRNACanProMet"]] <- dict_Vanderbilt[["BPRNACan"]]
dict_Vanderbilt[["BPRNACan3DProMet"]] <- dict_Vanderbilt[["BPRNACan"]]

dict_Vanderbilt[["CBSX_Fig2ab-NSCLC_PBMCs_scRNAseq_sigmatrix"]] <- list(`B cells` = c("B_cells"),
                                      CD4 = c("T.cells.CD4"),
                                      CD8 = c("T.cells.CD8"),
                                      `Endothelial cells` = c(),
                                      Cancer = c(),
                                      Fibroblasts = c(),
                                      `M1 Macrophages` = c(),
                                      `M2 macrophages` = c(),
                                      `Mast cells` = c(),
                                      `Mural cells` = c(),
                                      NK = c("NK"), # NKT.cells?
                                      `Plasma B cells` = c(),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c(),
                                      `T reg. cells` = c()) # Missing Monocytes and NKT.cells

dict_Vanderbilt[["CBSX_LM22"]] <- list(`B cells` = c("B.cells.memory", "B.cells.naive"),
                                      CD4 = c("T.cells.CD4.memory.activated", "T.cells.CD4.memory.resting", "T.cells.CD4.naive"),
                                      CD8 = c("T.cells.CD8", "T.cells.gamma.delta"),
                                      `Endothelial cells` = c(),
                                      Cancer = c(),
                                      Fibroblasts = c(),
                                      `M1 Macrophages` = c("Macrophages.M1"),
                                      `M2 macrophages` = c("Macrophages.M2"),
                                      `Mast cells` = c("Mast.cells.activated", "Mast.cells.resting"),
                                      `Mural cells` = c(),
                                      NK = c("NK.cells.activated", "NK.cells.resting"), 
                                      `Plasma B cells` = c("Plasma.cells"),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c("T.cells.follicular.helper"),
                                      `T reg. cells` = c("T.cells.regulatory..Tregs.")) # Missing Dendritic.cells.activated, Dendritic.cells.resting, Eosinophils, Macrophages.M0, , Monocytes, Neutrophils

dict_Vanderbilt[["CBSX_scRNA-Seq_melanoma_Tirosh_sigmatrix_SuppFig_3-b"]] <- list(`B cells` = c("B.cells"),
                                      CD4 = c("T.cells.CD4"),
                                      CD8 = c("T.cells.CD8"),
                                      `Endothelial cells` = c("Endothelial.cells"),
                                      Cancer = c("Malignant"),
                                      Fibroblasts = c("CAF"),
                                      `M1 Macrophages` = c(),
                                      `M2 macrophages` = c(),
                                      `Mast cells` = c(),
                                      `Mural cells` = c(),
                                      NK = c("NK.cells"), 
                                      `Plasma B cells` = c(),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c(),
                                      `T reg. cells` = c()) # Missing Macrophages

list(`B cells` = c(),
                                      CD4 = c(),
                                      CD8 = c(),
                                      `Endothelial cells` = c(),
                                      Cancer = c(),
                                      Fibroblasts = c(),
                                      `M1 Macrophages` = c(),
                                      `M2 macrophages` = c(),
                                      `Mast cells` = c(),
                                      `Mural cells` = c(),
                                      NK = c(), 
                                      `Plasma B cells` = c(),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c(),
                                      `T reg. cells` = c())

dict_Vanderbilt[["CBSX_sigmatrix_HNSCC_Fig2cd"]] <- list(`B cells` = c("B.cells"),
                                      CD4 = c("T.cells.CD4"),
                                      CD8 = c("T.cells.CD8"),
                                      `Endothelial cells` = c(),
                                      Cancer = c("Malignant.cells"),
                                      Fibroblasts = c("CAFs"),
                                      `M1 Macrophages` = c(),
                                      `M2 macrophages` = c(),
                                      `Mast cells` = c("Mast.cells"),
                                      `Mural cells` = c(),
                                      NK = c(), 
                                      `Plasma B cells` = c(),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c(),
                                      `T reg. cells` = c()) # Missing Dendritic.cells Macrophages

dict_Vanderbilt[["Quantiseq"]] <- list(`B cells` = c("B_cell"),
                                      CD4 = c("T_cell_CD4+_(non-regulatory)"),
                                      CD8 = c("T_cell_CD8+"),
                                      `Endothelial cells` = c(),
                                      Cancer = c(),
                                      Fibroblasts = c(),
                                      `M1 Macrophages` = c("Macrophage_M1"),
                                      `M2 macrophages` = c("Macrophage_M2"),
                                      `Mast cells` = c(),
                                      `Mural cells` = c(),
                                      NK = c("NK_cell"), 
                                      `Plasma B cells` = c(),
                                      `T cells proliferative` = c(),
                                      `T helper cells` = c(),
                                      `T reg. cells` = c("T_cell_regulatory_(Tregs)")) # Missing Monocyte Myeloid_dendritic_cell Neutrophil uncharacterized_cell

dict_Vanderbilt[["Quantiseq"]] <- list(CD4 = c("T.cell.CD4+.(non-regulatory)"), # "T.cell.regulatory(Tregs)"
                                   CD8 = c("T.cell.CD8+"),
                                   NK = c("NK.cell"))

```{r}
corr <- benchmark_corrplots(deconv_data = deconv_Vanderbilt, true_estimates = truth_Vanderbilt, dict = dict_Vanderbilt, key = "Samples", title = "Vanderbilt scRNAseq")
corr
